---
output: html_document
---
RepData_PeerAssessment1
============================
## What is mean total number of steps taken per day?
### Part 1
```{r, echo=TRUE, eval=TRUE}
# Attaching required libraries
require("lattice")
require("dplyr")
```
#### Read in the activity.csv file
```{r, echo=TRUE, eval=TRUE}
activity<-read.csv("activity.csv", stringsAsFactors=F)
```
#### Convert "chr" forms of dates from the activity file to "Date" class
```{r}
dates<-as.POSIXlt(strftime(activity$date, format="", usetz=F))
dates<-as.Date(dates, format="%d%m%Y")
activity$date<-dates
```
#### Use the dplyr library to calculate the total steps per day
```{r}
steps_perDayTotal<-summarise(group_by(activity, date), sum(steps))
summary(steps_perDayTotal)
```
####
```{r, barchart, fig.width=6, fig.height=10, echo=FALSE}
print(barchart(`date`~`sum(steps)`, data=steps_perDayTotal, xlab="Total Steps", ylab="Date", scales=list(alternating='1', cex=c(0.9,1)), auto.key=T, horizontal=T, main="Total Steps per Day"))
```

#### Use the dplyr library to calculate the mean steps per day
```{r}
steps_perDayMean<-summarise(group_by(activity, date), mean(steps))
summary(steps_perDayMean)
head(steps_perDayMean)
```
#### Use the dplyr library to calculate the median steps per day
```{r}
steps_perDayMedian<-summarise(group_by(activity, date), median(steps))
summary(steps_perDayMedian)
head(steps_perDayMedian)
```
## What is the average daily activity pattern?
### Part 1
```{r}
cat("Hi")
```
##Inputing missing values
### Part 1
```{r}
activity<-dplyr:::mutate(activity, NAs = is.na(steps))
activity_NAs<-dplyr:::filter(activity, NAs==TRUE)%>%summarise(n())
summary(activity_NAs)
````